VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-06-27 snapshot_system/paper_background.tex
revise for ACM cloud

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
efficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

%Further, 
%this artificial dependence affects fault isolation since 
%the loss of all replicas of a data chunk can affect the otherwise unrelated
%VMs that share it.

Indeed, the greater the deduplication efficiency, the less
the data redundancy is maintained, and the greater the ``hidden'' data
dependence between VMs that must be hosted as if they were physically isolated.
%machine failures happen periodically 
%in a large-scale cloud and
%loss of a small number of shared data chunks can 
%cause the unavailability of snapshots for a large number of virtual machines.
Thus,
restricting deduplication to be local to each VM (but across snapshots) 
aligns the units of storage with the isolation properties guaranteed by the
cloud at the possible expense of storage efficiency.
%Localizing the impact of deduplication can increase fault isolation and resilience.
%Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
%discouraged.

@@

For the remainder of this paper, 
we term the traditional deduplication approach {\em VM-oblivious} (VO)
because it manages duplicate data chunks without consideration of VMs.
With the above  considerations in mind, we propose a 
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.

In designing a VC deduplication solution, we have considered and adopted some of
the following previously-developed techniques.
1) {\em Changed Block Tracking}. 
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized to the snapshot
history for each VM.
%We are seeking for a tradeoff since 
%global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
1) {\em Stateless  Data Routing}.
One approach to scalable duplicate comparison is to use content-based hash
partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index for Memory Parsimony}. 
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation 
and discuss snapshot deletion support, then present an implementation design.

>>>>

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
In a converged setting with source-side deduplication, the resources that are used to implement snapshot
backup and deduplication are the same resources that must support cloud-hosted
VMs.  Thus the backup service colocated with 
other cloud services has to minimize its resource impact.  
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
efficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}.

\comments{
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
}
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

Even with these optimization and approximation techniques, resource usage for deduplication
is still extensive for a shared cloud cluster.  For example, the
sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011},
uses 5GB memory space for 100TB of raw data even the compressed data size
is about 5TB. It is fairly typical for
a physical machine with terabytes of disk storage and tens of gigabytes of memory
to support  more than 20 VMs and 10 snapshots per VM and can require  over 100TB of raw space
even the actual storage space need is about 5TB assuming 20:1 compression ratio with deduplication.
It is not easy  to adopt the sampled index method  for a distributed architecture
and every deduplication request may access a remote machine for index lookup and the overall overhead of access latency 
for all requests can be significant.  In general
with a cluster of machines in a cloiud cluster, there will be more memory requirement per machine such as a 
fingerprint cache, for cross-machine fingerprint comparison~\cite{bottleneck08,Dong2011} and a deduplication
task often needs several gigabytpes of memory, which impacts other cloud services sharing the omputing resource.

Whiel deduplication reduces storage cost, it creates an artitical data dependence among VMs: when a shared data chunk
is lost in the storage, all VMs that refer such chunks after deduplication face a data loss.
This issue is identified in ~\cite{Baghat2006} which proposes to use more replication for data shared by more files.
It is complex to implement this idea in VM snapshot backup. 
This is because the file system block  size in a distributed file system such as  
the Hadoop file system and the Google file system~\cite{googlefs03}
is large (e.g.  64MB) so as to optimize file system performance
for big data. Since  deduplication implementations~\cite{Guo2011,extreme_binning09,bottleneck08,Jin2009,Dong2011}
 typically use smaller chunk sizes (e.g. 4K bytes),  a large file block contains many small chunks with different 
degree of sharing and  these file blocks would have to be replicated.

Cross-VM chunk sharing also complicates snapshot deletion which occurs frequently when snapshots expire regularly
and reference counting is required to identify chunks that are no longer used. 
The mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  is effective for
deletion, but still carries a significant cost, especially in a distributed 
setting.  
%to count if a data chunk is still shared by other snapshots.

%still significant
%especailly for a distributed setting.

\comments{
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

Indeed, the greater the deduplication efficiency, the less
the data redundancy is maintained, and the greater the ``hidden'' data
dependence between VMs that must be hosted as if they were physically isolated.
Thus,
restricting deduplication to be local to each VM (but across snapshots) 
aligns the units of storage with the isolation properties guaranteed by the
cloud at the possible expense of storage efficiency.
%Localizing the impact of deduplication can increase fault isolation and resilience.
%Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
%discouraged.

@@

}

\comments{
With the above  considerations in mind, we propose a 
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.  
This compares the traditional deduplication approach {\em VM-oblivious} (VO)
which manages duplicate data chunks without consideration of VMs.
We discuss our key ideas and design in the next section.
%Our key idea is to simplify deduplication management and 
}
\comments{
In designing a VC deduplication solution, we have considered and adopted some of
the following previously-developed techniques.
1) {\em Changed Block Tracking}. 
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized to the snapshot
history for each VM.
%We are seeking for a tradeoff since 
%global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
1) {\em Stateless  Data Routing}.
One approach to scalable duplicate comparison is to use content-based hash
partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index for Memory Parsimony}. 
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation 
and discuss snapshot deletion support, then present an implementation design. 
}

### 2014-06-27 snapshot_system/paper_dedup.tex
update graph
revise for ACM cloud

%\subsection{VM centric  strategies}
%\label{sect:vc-strategies}

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search, global deduplication using popular chunks, and 
VM-centric file system block management. 
%\begin{itemize}
%\item

\textbf{VM-local duplicate search optimization} --
We use the standard version-based detection~\cite{Clements2009,Vrable2009}
to identify changed content with dirty bits in a coarse grain segment level.
%In our implementation 
%the segment size is 2MB
%and the device driver is extended to support tracking changed segments using a dirty bitmap. 
The reason to choose a coarse grain level is that 
since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in 
memory and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. 
%% this method is so widely-used and easy to understand, everyone include Pure Storage has it, I feel no need to explain further.
%{
%For example,
%the VMWare hypervisor has an API to let external backup applications know 
%the changed areas since last backup. 
%The Microsoft SDK provides an API that allows external applications to monitor 
%the VM's I/O traffic and implement such changed block tracking feature.
%}

Since the best deduplication uses a non-uniform chunk size 
with an average of 4KB or 8KB~\cite{Jin2009},
we conduct additional local similarity guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segment recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata.
By limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

%\item 
\textbf{Global deduplication using popular chunks} --
This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is that the local deduplication has removed most of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
This number can be computed periodically, e.g., on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance.

@@

@@

%\item 
\textbf{VM-centric file system block management} --
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. 
%We manage this accumulation process using a log-structured storage scheme built
%on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

%The system allows all machines conduct backups in parallel, and each machine
%conducts the backup of one VM at a time, and thus only requires a write buffer for one VM.
%\end{itemize}

 \begin{figure}
 \centering
  \includegraphics[width=2.5in]{figures/log-log-disk}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

>>>>

With the considerations discussed in the previous section, we propose a
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.
This compares the traditional deduplication approach {\em VM-oblivious} (VO)
which manages duplicate data chunks without consideration of VMs.
Section~\ref{sect:vc-strategies} discusses  our key ideas and design for duplicate detection.
Section~\ref{sect:delete} presents  a simplified snapshot deletion with  the VM-centric approach.

\subsection{VM centric  strategies}
\label{sect:vc-strategies}

We describe our overall deduplication approach as consisting of three
complementary strategies: VM-local
duplicate search, global deduplication using popular chunks, and 
VM-centric file system block management. 
\begin{itemize}
\item 
\textbf{Cross-VM global deduplication using popular chunks} --
We seperate the deduplication within a VM and cross VMs.
This is because global deduplication that detects the apperance of a chunk 
in any VM requires a substantial resource for fingerprint comparison.
%This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is for those chunks that appear in different VMs, top $k$ most popular items
dominate the apperance. 
 \begin{figure}
 \centering
  \includegraphics[width=2.5in]{figures/zipf_count_rank.pdf}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

@@

%when restricting fingerprint comparison within a VM uses
%only a small amount of resource. 
%ehat the local deduplication has removed most of the duplicates.
%There are fewer deduplication opportunities across VMs while the memory and network
%consumption for global comparison is more expensive.
Thus our approximation is that the global cross-VM fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
i.e., the number of copies of the chunk after local deduplication.
This number can be computed periodically, e.g., on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance.

@@

\item 
\textbf{VM-specific duplicate search optimization} --
Fingerprint lookup within a VM does not require a signficant  amount of memory resource.
While cross-VM deduplication is simplified, we intend to optimize the VM-sepcific deduplication 
as much as possible.

We start with the standard version-based detection~\cite{Clements2009,Vrable2009}
to identify changed content with dirty bits in a coarse grain segment level.
%In our implementation 
%the segment size is 2MB
%and the device driver is extended to support tracking changed segments using a dirty bitmap. 
The reason to choose a coarse grain level is that 
since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in 
memory and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. 
%% this method is so widely-used and easy to understand, everyone include Pure Storage has it, I feel no need to explain further.
%{
%For example,
%the VMWare hypervisor has an API to let external backup applications know 
%the changed areas since last backup. 
%The Microsoft SDK provides an API that allows external applications to monitor 
%the VM's I/O traffic and implement such changed block tracking feature.
%}

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segment recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata.
By limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

\item 
\textbf{VM-centric file system block management} --
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. 
%We manage this accumulation process using a log-structured storage scheme built
%on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

%The system allows all machines conduct backups in parallel, and each machine
%conducts the backup of one VM at a time, and thus only requires a write buffer for one VM.
\end{itemize}

### 2014-06-27 snapshot_system/paper_intro.tex
revise for ACM cloud

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot
and are shared among VMs,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

%Without deduplication,
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out undeduplicated backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).

This paper considers a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource with them.
We describe methodologies to address the following two challenges: 
1) simple deduplication and snapshot deletion with competitive efficiency and 
a small resource usage without affecting other collocated cloud services;
2) improved fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception. 
% so that users stop incurring storage charges immediately after snapshot deletion.
%\end{itemize}

% low-cost VM-centric snapshot backup approach that is co-located within 
% the existing converged storage architecture. Our solution leverages the scalability of the underlying
% distributed file system and provides global data deduplication in a resource friendly manner.

%For example, the Aliyun cloud, which is  the largest cloud service provider by Alibaba in China, 
%automatically conducts  the backup of virtual disk images to all active users every day.
%The cost of supporting a large number of concurrent backup streams is high
%because of the huge storage demand and the use of deduplication

%Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
%can effectively identify and remove content duplicates among snapshots, 
%but such a solution can be expensive. There is also a large amount of 
%network traffic to transfer  data from the host machines to the backup facility
%before duplicates are removed.

Version-based detection has been used to identify file blocks that have not
changed across snapshots so their redundant storage can be
avoided~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%the previous version of the
%snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}, 
Alternatively, techniques that compare block ``fingerprints'' to
identify duplicates
that exist among all files regardless of their 
origin~\cite{Guo2011,Dong2011,extreme_binning09} have
also gained in popularity.  
One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  
%The problem is particularly acute for VM
%snapshots where the purpose of the snapshot is precisely to create redundant
%data copies so that VMs are more fault resilient.  
Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiency that deduplication makes possible.
The previous work on deduplication has not considered the impact of 
duplicate sharing on  fault tolerance.
%Because
%of highly repetitive content in snapshots from different VMs, many data chunks
%are shared by virtual machines.  Failure of a few shared data chunks can have
%a broad effect and snapshots of these virtual machines could be affected.  The
%previous work in deduplication focuses on the efficiency and approximation of
%fingerprint comparison, and has not addressed fault tolerance issues  together
%with deduplication.

%This paper explores the storage and
%performance efficiencies that such a {\em VM-centric} 
%combination of local and global
%techniques can achieve.  
We term our approach {\em VM-centric} because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud environments.
Our work focuses on sharing only ``popular'' data blocks 
(through deduplication and limited replication)
across 
virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
%, and to propose a VM-centric approach that localizes
%deduplication as much as possible and restricts global deduplication only to a
%limited set of most popular blocks.  
%In our method, local deduplication uses
%similarity-guided elimination to improve the deduplication coverage.  
Since the underlying storage file system block size is typically bigger than the average data chunk size
used for deduplication,  packaging
content from different VMs into the same file block  creates a ``false'' sharing that 
affects fault isolation. Thus we choose to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and
%Thus our strategy is to packages data chunks from the same VM into a file
%system block as much as possible to improve fault isolation.  
%Deduplicating
%data locally within a VM also reduces the need to maintain a global index of
%chunks that are shared among VMs, thereby increasing the scaling efficiency of
%the snapshot system.  
The cost associated with localization, however, is in 
the lost
storage efficiency that cross-VM deduplication makes possible.  To recover
some of this efficiency we allow a small number of the most popular data
blocks to be deduplicated (and then replicated) across VMs.

>>>>

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot
and are shared among VMs,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.

%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantic Netbean})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
But it is memory-intensive to compare a large number of fingerprints and identify
duplicates, even optimization or approximation techniques are developed~\cite{Guo2011,Dong2011,extreme_binning09}. 
When deleting unused snapshots, a mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  
is effective while it still carries a significant cost, especially in a distributed
setting.  Another side effect of deduplication is the possible loss of failure
resilience~\cite{Bhagwat2006}.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out undeduplicated backup data wastes huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

This paper considers a backup service for converged cloud architectures
that is co-located with user VMs, sharing the cloud computer, network, and
storage resource with them.  We propose a VM-centric approach that
simplifies deduplication and snapshot deletion.
The key benefit is that while maintaining  a competitive deduplication efficiency,
the scheme uses a small resource usage without affecting other collocated cloud services.
\comments{
2) improved fault isolation for the sharing of deduplicated data,
given that node failures in converged architecture are the norm rather than
exception.

Version-based detection has been used to identify file blocks that have not
changed across snapshots so their redundant storage can be
avoided~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%the previous version of the
%snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}, 
Alternatively, techniques that compare block ``fingerprints'' to
identify duplicates
that exist among all files regardless of their 
origin~\cite{Guo2011,Dong2011,extreme_binning09} have
also gained in popularity.  
One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  
Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiency that deduplication makes possible.
The previous work on deduplication has not considered the impact of 
duplicate sharing on  fault tolerance.
} 
We term our approach {\em VM-centric} because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
Our work focuses on a tradeoff that allows the sharing of only ``popular'' data blocks 
across virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
\comments{
We detail the storage, performance, and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud environments.

%, and to propose a VM-centric approach that localizes
%deduplication as much as possible and restricts global deduplication only to a
%limited set of most popular blocks.  
%In our method, local deduplication uses
%similarity-guided elimination to improve the deduplication coverage.  
Since the underlying storage file system block size is typically bigger than the average data chunk size
used for deduplication,  packaging
content from different VMs into the same file block  creates a ``false'' sharing that 
affects fault isolation. Thus we choose to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and
%Thus our strategy is to packages data chunks from the same VM into a file
%system block as much as possible to improve fault isolation.  
%Deduplicating
%data locally within a VM also reduces the need to maintain a global index of
%chunks that are shared among VMs, thereby increasing the scaling efficiency of
%the snapshot system.  
The cost associated with localization, however, is in 
the lost
storage efficiency that cross-VM deduplication makes possible.  To recover
some of this efficiency we allow a small number of the most popular data
blocks to be deduplicated (and then replicated) across VMs.  
}
